Historical Soil Organic Carbon Budget

### 2020-07-14 SOCbudgetPaper/methods.Rmd

#### Crop Residue Production
Crop residue production and management is based on a revised methodology of ([@bodirsky2012]) and key aspects are explained here given its central role for soil carbon modelling. Starting from crop production ($CP$) estimates of the harvested organs <!-- organs very odd here, and I'm not sure what the unit being measured/estimated is. Is it just crop yields? "Starting from respective crop yields ($CP$) and harvested area ($CA$)..." maybe simplify something like that  --> and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest indices ($HI$) and root:shoot ratios ($RS$) as follows

@@

##### Burned Residues
A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country. Following [@smil1999]) we assume 25% burn share for low-income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr}$ and linearly interpolate shares for all middle-income countries depending on their per-capita income. Depending on the crop group 80–90% of the residue carbon burned on the fields are lost within the combustion process ([@ipcc_2006_2006]).

##### Residue Usage
We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \@ref(append:Table_kcr2kres)). Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see [@weindl]), taking available AGR biomass as well as livestock productivity into account. We estimate, for low-income countries ($<\,1000\,\tfrac{USD}{yr}$) per capita,  a material-use share for straw residues of 5% and a fuel-share of 10% for all used residues groups in low income countries ($<\,1000\,\tfrac{USD}{yr}$). For high-income countries, ($>\,10000\,\tfrac{USD}{yr}$ no withdrawal for material or fuel use is assumend, and middle-income countries use shares are linearly interpolated based on per-capita income. The remaining AGR as well as all BGR are assumend to be recycled to the soil. We limit high recycling shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.

@@

#### Livestock Distribution
To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from by the approach of [@gilbert], and uses feed basket assumptions based on a revised methodology from @weindl. We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
For ruminants, we assume that livestock is located where the production of feed occurs. We distingush between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feeds, which we consider to be consumend in intensive systems.
For poultry, egg and monogastric meat production we use the per-capita income of the country to divide between intensive and extensive production systems. For low-income countries (<1000 USD/yr), we assume extensive production systems. We locate them according to the share of built-up areas based on the idea that these animals are held in subsistence or small-holder farming systems with a high labour per animal ratio. Intensive production associated with higher income countries, is distributed within a country using again share of crop-based feed production,  assuming that feed availability is the most driving factor for livestock location.

#### Manure Excretion, Storage and Recycling
Manure production and management is based on a revised methodology of ([@bodirsky2012]) and is presented here due to its central role in soil carbon modelling. Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets ([@weindl]), assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted. Carbon in excreted manure is estimated by applying fixed C:N ratios (given by [(@ipcc_2019_2019)]). 
Depending on the feed system we assume manure to be handled in four different ways:
All manure orginated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel. Manure fuel shares are estimated using IPCC default values (@ippc_2006_2006). 
Whereas for low-income countries (<1000 USD/yr), we adopt a share of 25% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries ($>\,10000\,\tfrac{USD}{yr}$; middle-income countries see linearly interpolated shares depending on their per-capita income.
For all other feed items we assume the manure to be stored in animal waste management systems associated with livestock housing. 
To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems. Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled. For manure stored in different animal waste management systems we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in [@bodirsky2012] (see \@ref(append:Table_clossAWMS))). We limit high application shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.

@@

We use irrigation area shares to modify the water effect on decomposition by weighting the irrigated and rainfed water factors based on these shares. The LUH2v2 data set provides irrigated fractions for their cropland subcategories. We apply aggregated area shares, leading to the same impact of irrigation on all of our crop groups. Moreover we assume the irrigation effect to be present for all months of a year, that has been marked as irrigated. <!-- assuming that irrigation effect happens 12 months? or just the months "marked as irrigated" ? this term confusing -->

>>>>

#### Crop Residue Production
Crop residue production and management is based on a revised methodology of ([@bodirsky2012]) and key aspects are explained here given its central role for soil carbon modelling. Starting from harvested crop production ($CP$) estimates and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest indices ($HI$) and root:shoot ratios ($RS$) as follows

@@

##### Burned Residues
A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country. Following [@smil1999]) we assume 25% burn share for low-income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income. Depending on the crop group 80–90% of the residue carbon burned on the fields are lost within the combustion process ([@ipcc_2006_2006]).

##### Residue Usage
We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \@ref(append:Table_kcr2kres)). Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see [@weindl]), taking available AGR biomass as well as livestock productivity into account. We estimate, for low-income countries,  a material-use share for straw residues of 5% and a fuel-share of 10% for all used residues groups in low-income countries. For high-income countries, no withdrawal for material or fuel use is assumend, and middle-income countries use shares are linearly interpolated based on per-capita income. The remaining AGR as well as all BGR are assumend to be recycled to the soil. We limit high recycling shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.

@@

#### Livestock Distribution
To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from by the approach of [@gilbert], and uses feed basket assumptions based on a revised methodology from @weindl. We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
For ruminants, we assume that livestock is located where the production of feed occurs. We distingush between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feeds, which we consider to be consumend in intensive systems.
For poultry, egg and monogastric meat production we use the per-capita income of the country to divide between intensive and extensive production systems. For low-income countries, we assume extensive production systems. We locate them according to the share of built-up areas based on the idea that these animals are held in subsistence or small-holder farming systems with a high labour per animal ratio. Intensive production associated with higher income countries, is distributed within a country using again share of crop-based feed production,  assuming that feed availability is the most driving factor for livestock location.

#### Manure Excretion, Storage and Recycling
Manure production and management is based on a revised methodology of ([@bodirsky2012]) and is presented here due to its central role in soil carbon modelling. Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets ([@weindl]), assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted. Carbon in excreted manure is estimated by applying fixed C:N ratios (given by [(@ipcc_2019_2019)]). 
Depending on the feed system we assume manure to be handled in four different ways:
All manure orginated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel. Manure fuel shares are estimated using IPCC default values (@ippc_2006_2006). 
Whereas for low-income countries, we adopt a share of 25% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
For all other feed items we assume the manure to be stored in animal waste management systems associated with livestock housing. 
To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems. Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled. For manure stored in different animal waste management systems we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in [@bodirsky2012] (see \@ref(append:Table_clossAWMS))). We limit high application shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.

@@

We use irrigation area shares to modify the water effect on decomposition by weighting the irrigated and rainfed water factors based on these shares. The LUH2v2 data set provides irrigated fractions for their cropland subcategories. We apply aggregated area shares, leading to the same impact of irrigation on all of our crop groups. Moreover we assume the irrigation effect to be present for all 12 months of a year in grid cells, that has been marked as irrigated.
